---
title: "HW7_SDS315"
author: "Nidhi Ilanthalaivan"
date: "2025-04-06"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(MatchIt)
library(tidyverse)
```
## Problem 1: Armfolding

### Part A: Load and examine the data

```{r echo=FALSE, warning=FALSE, message=FALSE}
setwd("/Users/nidhi/Downloads/")
armfold <- read.csv("armfold.csv")

table(armfold$Sex)

male_left <- sum(armfold$LonR_fold == 1 & armfold$Sex == "Male")
male_total <- sum(armfold$Sex == "Male")
prop_male_left <- male_left / male_total

female_left <- sum(armfold$LonR_fold == 1 & armfold$Sex == "Female")
female_total <- sum(armfold$Sex == "Female")
prop_female_left <- female_left / female_total

prop_male_left
prop_female_left

```
### Part B: Observed Difference in Proportions

```{r echo=FALSE, warning=FALSE, message=FALSE}
observed_difference <- prop_male_left - prop_female_left
cat("Observed Difference", observed_difference)

```

### Part C: 95% Confidence Interval - Manual


```{r echo=FALSE, warning=FALSE, message=FALSE}

SE <- sqrt((prop_male_left * (1 - prop_male_left) / male_total) + 
           (prop_female_left * (1 - prop_female_left) / female_total))


z_val <- 1.96  
lower <- observed_difference - z_val * SE
upper <- observed_difference + z_val * SE

cat("95% CI: [", lower, ",", upper, "]\n\n")

```
### Part C: 95% Confidence Interval - RStudio

```{r echo=FALSE, warning=FALSE, message=FALSE}

prop_test_result <- prop.test(c(male_left, female_left), c(male_total, female_total), conf.level = 0.95)
prop_test_result$conf.int

```
Both the manual results and the RStudio results are about the same amount.

### Part D: Interpret Confidence Interval

If we were to repeat this study many times, then we would expect that the true population parameter will lie within this interval approxiately 95% of the time. 


### Part E: Standard Error Representation

The standard error represents the estimated variability of the difference in proportions across different random samples from the population. The value of 0.067 represents the estimated variability in the difference of proportions of males and females who fold their arms with the left arm on top if we were to draw different random samples from the same population of students. Considering that this standard error is smaller, this suggests that the sample difference in proportions is more precise. The sample proportions will vary because we are drawing different students from the population each time but the true population proportions and the sampling procedure remains the exact same.

### Part F: Sampling Distribution

The sampling distribution refers to the distribution of the difference in sample proportions (males minus females) that we can get if we took many random samples from the population and calculate the difference in proportions (left arm on top for males and females) in each sample. The sampling distribution is the theoretical distribution of the observed differences in sample proportions across multiple random samples. The difference in sample proportions will change because each sample can have different proportions of males and females who fold their arms with the left arm on top. But the true population proportions and the sampling process stays the same. The variability in the sampling distribution is represented by the standard error.

### Part G: Central Limit Theorem

The mathematical result that justifies using a normal distribution to approximate the sampling distribution of the difference in sample proportions is the Central Limit Theorem. The theorem explains that if we take a large number of random samples from any population, the sampling distribution of the sample proportions will approach a normal distribution as long as the sample sizes are large enough. As long as that's true for both the female populations and male populations in our data, we are confident that the difference in proportions follows a normal distribution closely enough for us to use things like confidence intervals and hypothesis tests.

### Part H: Verifying Confidence Interval

If someone claimed that there is no sex difference in arm folding, I would look at the intervals from the 95% confidence interval which ranges from -0.01 to 0.30. Considering that this interval includes zero, we cannot rule out the chance that there is no true difference between males and females in the population. While the observed data suggests that males may be more likely to fold their left arm on top, this evidence is not statistically significant at the 95% confidence interval. It is concludable that the observed difference is due to random variation in the sample. We fail to reject the hypothesis that there is no sex-based difference in arm folding activity. 

### Part I: Would Confidence Interval be Different?

Yes, the confidence interval would vary across different random samples students. This variation occurs because each sample may yield slightly different proportions due to natural random differences in who is selected. As a result, the observed difference in proportions and the resulting confidence interval will not be identical every time. However, if we were to repeat this experiment many times and calculate a 95% confidence interval each time, we would expect about 95% of those intervals to contain the true population difference in proportions. A 95% confidence level tells us how well the method performs in the long run, instead of guaranteeing the accuracy of a single interval.

## Problem 2

### Part A

```{r echo=FALSE, warning=FALSE, message=FALSE}

df <- read.csv("turnout.csv")

gotv_yes <- subset(df, GOTV_call == 1)
prop_gotv_yes <- mean(gotv_yes$voted1998)
cat("proportion of GOTV call recipients who voted in 1998:", prop_gotv_yes, "\n")


gotv_no <- subset(df, GOTV_call == 0)
prop_gotv_no <- mean(gotv_no$voted1998)
cat("proportion of non-recipients who voted in 1998:", prop_gotv_no, "\n")

n1 <- nrow(gotv_yes)
n2 <- nrow(gotv_no)
p1 <- prop_gotv_yes
p2 <- prop_gotv_no

se_diff <- sqrt((p1*(1 - p1)/n1) + (p2*(1 - p2)/n2))
z_star <- 1.96  
margin <- z_star * se_diff


diff_prop <- p1 - p2
ci_lower <- diff_prop - margin
ci_upper <- diff_prop + margin

cat("Difference in proportions (GOTV - No GOTV):", diff_prop, "\n")
cat("95% Confidence Interval: [", ci_lower, ",", ci_upper, "]\n")


```
### Part B: Variables are Confounders or Not?

```{r echo=FALSE, warning=FALSE, message=FALSE}


gotv_yes <- df[df$GOTV_call == 1, ]
gotv_no <- df[df$GOTV_call == 0, ]

mean_yes_1996 <- mean(gotv_yes$voted1996)
mean_no_1996 <- mean(gotv_no$voted1996)
diff_1996 <- mean_yes_1996 - mean_no_1996

se_1996 <- sqrt((var(gotv_yes$voted1996)/nrow(gotv_yes)) + (var(gotv_no$voted1996)/nrow(gotv_no)))
lower_1996 <- diff_1996 - 1.96 * se_1996
upper_1996 <- diff_1996 + 1.96 * se_1996

cat("voted1996:\n")
cat("Mean (GOTV = 1):", mean_yes_1996, "\n")
cat("Mean (GOTV = 0):", mean_no_1996, "\n")
cat("Difference:", diff_1996, "\n")
cat("95% CI: [", lower_1996, ",", upper_1996, "]\n\n")


mean_yes_age <- mean(gotv_yes$AGE)
mean_no_age <- mean(gotv_no$AGE)
diff_age <- mean_yes_age - mean_no_age


se_age <- sqrt((var(gotv_yes$AGE)/nrow(gotv_yes)) + (var(gotv_no$AGE)/nrow(gotv_no)))
lower_age <- diff_age - 1.96 * se_age
upper_age <- diff_age + 1.96 * se_age

cat("AGE:\n")
cat("Mean (GOTV = 1):", mean_yes_age, "\n")
cat("Mean (GOTV = 0):", mean_no_age, "\n")
cat("Difference:", diff_age, "\n")
cat("95% CI: [", lower_age, ",", upper_age, "]\n\n")

mean_yes_party <- mean(gotv_yes$MAJORPTY)
mean_no_party <- mean(gotv_no$MAJORPTY)
diff_party <- mean_yes_party - mean_no_party


se_party <- sqrt((var(gotv_yes$MAJORPTY)/nrow(gotv_yes)) + (var(gotv_no$MAJORPTY)/nrow(gotv_no)))
lower_party <- diff_party - 1.96 * se_party
upper_party <- diff_party + 1.96 * se_party

cat("MAJORPTY:\n")
cat("Mean (GOTV = 1):", mean_yes_party, "\n")
cat("Mean (GOTV = 0):", mean_no_party, "\n")
cat("Difference:", diff_party, "\n")
cat("95% CI: [", lower_party, ",", upper_party, "]\n")


```
 
To determine whether voted1996, AGE, and MAJORPTY are confounders, we compare their means between individuals who received a GOTV call and those who did not. If a variable both predicts receiving a GOTV call and is associated with voting in 1998, it may confound the relationship we are studying.

For voted1996, the average among those who received a GOTV call is 0.713, compared to 0.531 for those who did not. The difference of 0.182 has a 95% confidence interval of [0.124, 0.239], indicating a statistically significant difference in prior voting behavior. This means that people who voted previously were more likely to receive a GOTV call and are also more likely to vote again, making this a clear confounder.

For AGE, the average age of those who received a GOTV call is 58.31, and for those who did not, it is 49.43. The difference of 8.88 years is substantial, with a 95% confidence interval of [6.38, 11.38]. This means older individuals were more likely to receive a call. Age is typically associated with higher turnout, making AGE another confounder.

For MAJORPTY, the proportion registered with a major party is 0.802 among GOTV call recipients and 0.745 among non-recipients, with a difference of 0.057. The 95% confidence interval for this difference is [0.006, 0.107]. This suggests a significant association. Since major party affiliation could influence both call targeting and voting behavior, this variable may act as a confounder.

In summary, all three variables show significant differences across treatment groups and are confounders that must be accounted for to estimate the true effect of the GOTV call.

### Part C: 

```{r echo=FALSE, warning=FALSE, message=FALSE}


match_results <- matchit(GOTV_call ~ voted1996 + AGE + MAJORPTY,
                         data = df,
                         method = "nearest",
                         ratio = 5)


matched_data <- match.data(match_results)

matched_data %>% group_by(GOTV_call) %>%
  summarise(voted1996 = mean(voted1996),
            AGE = mean(AGE),
            MAJORPTY = mean(MAJORPTY))

t.test(voted1996 ~ GOTV_call, data = matched_data)
t.test(AGE ~ GOTV_call, data = matched_data)
t.test(MAJORPTY ~ GOTV_call, data = matched_data)

matched_data %>%
  group_by(GOTV_call) %>%
  summarise(voted1998 = mean(voted1998))


t.test(voted1998 ~ GOTV_call, data = matched_data)


```
After performing matching to balance the confounding variables, we re-analyzed the effect of receiving a GOTV call on voting in the 1998 Congressional election. The results show that among the matched sample, 64.8% of individuals who received a GOTV call voted, compared to 56.9% of those who did not receive a call. A Welch two-sample t-test shows that this difference is statistically significant, with a p-value of approximately 0.0198. The 95% confidence interval for the difference in voting proportions is approximately [-0.14, -0.01].This suggests that receiving a GOTV call is associated with an increase in the likelihood of voting. Since the confidence interval does not contain zero, we can conclude that the GOTV call had a positive effect on voter turnout in the 1998 election.

