---
title: "SDS315 Homework 8"
author: "Nidhi Ilanthalaivan (ni2649) - SDS 315 UT Austin"
date: "https://github.com/nidhi-ilanthalaivan/SDS-315/tree/main"
output:
  pdf_document: 
    latex_engine: xelatex
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=FALSE, warning=FALSE, message=FALSE}
library(tidyverse)
library(boot)

```
# Problem 1: regression warm up

## A) What creatine clearance rate should we expect for a 55-year-old?

```{r echo=FALSE, warning=FALSE, message=FALSE}
setwd("/Users/nidhi/Downloads/")
creatine <- read.csv("creatinine.csv")
model <- lm(creatclear ~ age, data = creatine)
summary(model)



```
**Regression Equation: 147.813 + (-0.61982 x age)**

## A) What creatine clearance rate should we expect for a 55-year-old?

creatclear = 147.813 + (-0.61982 x 55)
creatclear = 147.813 - 34.0901 = 113.72 mL/min

I plugged age = 55 into the regression equaton to predict the expected creatine clearance rate.

## B) How does creatine clearance rate change with age

The slope is **-0.61982 mL/min** per year. The slope of the regression line tells us that for each additional year of age, creatine clearance decreased by approximately 0.62 mL/min. 

## C) Whose creatine clearance rate is healthier (higher) for their age: a 40-year-old with a rate of 60-year-old with a rate of 112? 

40-year-old with observed clearance of 135:
creatclear of 40 = 147.813 + (- 0.61982 x 40) 
creatclear of 40 = 147.813 - 24.7928 = 123.0202
residual of 40 = 135 - 123.0202 ~ **11.98**

60-year-old with observed clearance of 112:
creatclear of 60 = 147.813 + (-0.61982 x 60) 
creatclear of 60 = 147.813 - 37.1892 = 110.6238
residual of 60 = 112 - 110.6238 = **1.38**

To determine who is healthier for their age, we compare each person's creatinine clearance rate to what would be expected based on their age using the regression model:

For a 40-year-old, the predicted clearance rate is 123.02 mL/min, and for a 60-year-old, it is 110.62 mL/min. The actual clearance values are 135 and 112, respectively. We calculate residuals by subtracting the predicted value from the actual value. The 40-year-old has a residual of 135 − 123.02 = 11.98, while the 60-year-old has a residual of 112 − 110.62 = 1.38. Since the 40-year-old has a larger positive residual, they are performing much better than expected for their age. **Therefore, the 40-year-old is healthier relative to their age group.**

# Problem 2: Modeling disease growth

```{r echo=FALSE, warning=FALSE, message=FALSE}

covid <- read.csv("covid.csv")

# Log-transform deaths
covid <- covid %>%
  filter(deaths > 0) %>%
  mutate(log_deaths = log(deaths))

# Separate by country
italy <- filter(covid, country == "Italy")
spain <- filter(covid, country == "Spain")

# Fit linear models on log scale
fit_italy <- lm(log_deaths ~ days_since_first_death, data = italy)
fit_spain <- lm(log_deaths ~ days_since_first_death, data = spain)

summary(fit_italy)
summary(fit_spain)


```
```{r echo=FALSE, warning=FALSE, message=FALSE}

b_italy <- coef(fit_italy)[2]
b_spain <- coef(fit_spain)[2]

dt_italy <- log(2) / b_italy
dt_spain <- log(2) / b_spain

```

```{r echo=FALSE, warning=FALSE, message=FALSE}

boot_growth <- function(data, indices) {
  d <- data[indices, ]
  model <- lm(log_deaths ~ days_since_first_death, data = d)
  coef(model)[2]  
}

#italy
set.seed(123)
boot_italy <- boot(italy, boot_growth, R = 1000)
boot.ci(boot_italy, type = "perc")

#spain
set.seed(123)
boot_spain <- boot(spain, boot_growth, R = 1000)
boot.ci(boot_spain, type = "perc")

dt_italy_boot <- log(2) / boot_italy$t
dt_spain_boot <- log(2) / boot_spain$t
quantile(dt_italy_boot, c(0.025, 0.975))
quantile(dt_spain_boot, c(0.025, 0.975))




```

```{r echo=FALSE, warning=FALSE, message=FALSE}
ggplot(covid, aes(x = days_since_first_death, y = deaths, color = country)) +
  geom_line(linewidth = 1) +
  labs(
    title = "Daily COVID-19 Deaths Over Time",
    x = "Days Since First Death",
    y = "Deaths per Day",
    color = "Country"
  ) +
  theme_minimal()

```

To investigate early pandemic dynamics, we modeled COVID-19 deaths in Italy and Spain using exponential growth models. Specifically, we regressed the log of daily death counts on the number of days since each country’s first reported death.

For Italy, the estimated daily growth rate was 0.183 (SE = 0.0103), with a 95% bootstrap confidence interval of (0.158, 0.208). This implies that daily deaths were increasing by approximately 18.3% each day. The corresponding doubling time was 3.8 days, with a 95% CI of (3.33, 4.38). The model explained 89.5% of the variance in log deaths (R squared = 0.895), and the growth rate was highly statistically significant (p < 0.001).

In Spain, the estimated daily growth rate was even higher at 0.276 (SE = 0.0191), with a 95% bootstrap confidence interval of (0.233, 0.320). This translates to a faster doubling time of 2.5 days, with a 95% CI of (2.17, 2.97). The model fit was similarly strong (R squared = 0.889), and again the growth rate was highly significant (p < 0.001).

These results confirm that during the early stages of the outbreak, both countries experienced rapid exponential increases in deaths. Spain’s faster growth rate and shorter doubling time highlight a more aggressive trajectory. These findings emphasize the importance of early intervention during pandemic outbreaks.


# Problem 3: price elasticity of demand

```{r echo=FALSE, warning=FALSE, message=FALSE}
milk <- read_csv('milk.csv')
milk <- milk %>%
  mutate(log_price = log(price),
         log_sales = log(sales))
model <- lm(log_sales ~ log_price, data = milk)
summary(model)

```
```{r echo=FALSE, warning=FALSE, message=FALSE}

boot_fun <- function(data, indices) {
  sample_data <- data[indices, ]
  coef(lm(log_sales ~ log_price, data = sample_data))[2]
}

set.seed(123)
boot_out <- boot(data = milk, statistic = boot_fun, R = 1000)

boot.ci(boot_out, type = "perc")


```
To estimate the price elasticity of demand for milk, I used a log-log regression based on the model Q = K*P^B, which linearizes to log(Q) = log(K) + Blog(P). I regressed log(sales) on log(price) using data from the stated preference study. The estimated price elasticity of demand(B) is -1.62. This essentially means that a 1% increase in the price of milk leads to an approximate 1.62% decrease in quantity demanded. I then bootstrapped the regression 1000 times to construct a 95% confidence interval. The resulting 95% percentile-based confidence interval is (-1.77, -1.44), meaning that demand is elastic within this range.

