---
title: "HW2"
author: "Nidhi Ilanthalaivan"
date: "2025-01-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Beauty, or not, in the classroom


```{r, warning = FALSE, error = FALSE, echo = TRUE}

setwd("/Users/nidhi/Downloads/")
library(ggplot2)
library(dplyr)
profs <- read.csv("profs.csv")
```


### Part A: Histogram of Course Evaluation Scores

```{r, warning = FALSE, error = FALSE, echo = TRUE}
ggplot(profs, aes(x = eval)) + geom_histogram(binwidth = 0.25, fill = "steelblue", color = "black") + labs(title = "Distribution of Course Evaluation Scores", x = "Evaluation Score", y = "Count") + theme_minimal()

```



### Part B: Side-by-Side Boxplots for Native or Non-Native Professor's Evaluation Scores
```{r, warning = FALSE, error = FALSE, echo = TRUE}
ggplot(profs, aes(x = native, y = eval, fill = native)) + geom_boxplot() + labs(title = "Evaluation Scores by Native English Speaker Status", x = "Native English Speaker", y = "Evaluation Score") + scale_fill_manual(values = c("brown", "grey"), labels = c("Non-Native", "Native")) + theme_minimal()
```



### Part C: Faceted Histogram for Male and Female Instructors

```{r, warning = FALSE, error = FALSE, echo = TRUE}
ggplot(profs, aes(x= eval, fill = gender)) + geom_histogram(binwidth = 0.25, color = "black", alpha = 0.7) + facet_wrap(~gender, nrow = 2) + labs(title = "Evaluation Score Distributions by Gender", x = "Evaluation Score", y = "Count") + scale_fill_manual(values = c("darkgreen", "purple")) + theme_minimal()

```


### Part D: Scatterplot of Physical Attractiveness vs Evaluation Scores

```{r, warning = FALSE, error = FALSE, echo = TRUE}
ggplot(profs, aes(x = beauty, y = eval)) + geom_point(alpha = 0.6, color = "pink") + geom_smooth(method = "lm", color = "darkblue", se = FALSE) + labs(title = "Attractiveness vs Evaluation Scores", x = "Physical Attractiveness Rating", y = "Evaluation Score") + theme_minimal()


```


## Problem 2: Bike Sharing

### Plot A: Average Hourly Bike Rentals Line Graph

```{r, warning = FALSE, error = FALSE, echo = TRUE}
bike_data <- read.csv("bikeshare.csv")

hourly_avg <- bike_data %>% 
  group_by(hr) %>%
  summarise(avg_rentals = mean(total))

ggplot(hourly_avg, aes(x = hr, y = avg_rentals)) +geom_line(color = "blue", size = 1) + labs(title = "Average Hourly Bike Rentals", x = "Hour of the Day", y = "Average Total Rentals") + theme_minimal()

```


### Plot B: Hourly Rentals by Working Day

```{r, warning = FALSE, error = FALSE, echo = TRUE}

hourly_workingday_avg <- bike_data %>% group_by(hr, workingday) %>%
  summarise(avg_rentals = mean(total))

ggplot(hourly_workingday_avg, aes(x=hr, y = avg_rentals, group = workingday, color = factor(workingday))) +
  geom_line(size = 1) + 
  facet_wrap(~workingday, labeller = labeller(workingday = c("0" = "Non-Working Day", "1" = "Working Day"))) + labs(title = "Hourly Bike Rentals: Working vs. Non-Working Days",
x = "Hour of the Day", 
y = "Average Total Rentals",
color = "Day Type") + 
theme_minimal()



```


### Plot C: 9 AM Rentals by Weather and Working Day
```{r,warning = FALSE, error = FALSE, echo = TRUE}

bike_9am <- bike_data %>% 
  filter(hr == 9) %>% group_by(weathersit, workingday) %>% 
  summarise(avg_rentals = mean(total))

ggplot(bike_9am, aes(x = factor(weathersit), y = avg_rentals, fill = factor(weathersit))) + 
  geom_bar(stat = "identity") + 
  facet_wrap(~workingday, labeller = labeller(workingday = c("0" = "Non-Working Day", "1" = "Working Day"))) +
  labs(title = "9 AM Bike Rentals by Weather Condition", 
       x = "Weather Condition", 
       y = "Average Total Rentals",
       fill = "Weather Condition") +
  theme_minimal()


```

## Problem 3: Capital Metro UT Ridership

### Faceted Line Graph: Average Boardings by Hour of Day, faceted by day_of_week

```{r,warning = FALSE, error = FALSE, echo = TRUE}
capmetro_UT <- read.csv("capmetro_UT.csv")

capmetro_UT <- capmetro_UT %>% mutate(day_of_week = factor(day_of_week, levels = c("Mon", "Tue", "Wed", "Thu", "Fri", "Sat", "Sun")), month = factor(month, levels = c("Sep", "Oct", "Nov")))

avg_boardings <- capmetro_UT %>% 
  group_by(day_of_week, hour_of_day, month) %>%
  summarise(avg_boarding = mean(boarding, na.rm = TRUE), .groups = "drop")

plot1 <- ggplot(avg_boardings, aes(x = hour_of_day, y = avg_boarding, color = month)) +
  geom_line() + 
  facet_wrap(~ day_of_week) +
  labs(title = "Average Bus Boardings by Hour of Day, Day of Week, and Month", 
       x = "Hour of Day",
       y = "Average Boardings",
       color = "Month" +
  theme_minimal())

plot1

```

### Faceted Scatterplot Graph: Boardings vs Temperature, faceted by hour of the day

```{r,warning = FALSE, error = FALSE, echo = TRUE}

capmetro_UT %>%
  ggplot(aes(x = temperature, y = boarding, color = factor(weekend))) + 
  geom_point(alpha = 0.5) +
  facet_wrap(~hour_of_day) + 
  labs(
    title = "Boardings vs Temperature, Faceted by Hour", 
    x = "Temperature (F)",
    y = "Boardings",
    color = "Weekend"
  ) + 
  theme_minimal()

```

## Problem 4: Wrangling the Billboard Top 100

### Part A:
```{r, warning = FALSE, error = FALSE, echo = TRUE}
library(knitr)

billboard <- read.csv("billboard.csv")

top_songs_table <- billboard %>%
  group_by(performer, song) %>%
  summarise(week_count = n(), .groups = "drop") %>%
  arrange(desc(week_count)) %>%
  head(10)


kable(top_songs_table, caption = "Top 10 Songs with the Most Weeks on Billboard")

```
### Part B: Musical Diversity Over Time

```{r, warning = FALSE, error = FALSE, echo = TRUE}
library(ggplot2)

musical_diversity <- billboard %>%
  filter(year != 1958 & year != 2021) %>%
  group_by(year) %>%
  summarise(unique_songs = n_distinct(song), .groups = "drop")

ggplot(musical_diversity, aes(x = year, y = unique_songs)) + geom_line() + labs(
  title = "Musical Diversity of Billboard Top 100 Over Time",
  x = "Year",
  y = "Number of Unique Songs"
)



```

### Part C: Ten-Week Hits
```{r, warning = FALSE, error = FALSE, echo = TRUE}
library(forcats)


top_19_performers <- billboard %>%
  group_by(performer) %>%
  summarize(ten_week_hits = sum(weeks_on_chart >= 10)) %>%
  filter(ten_week_hits >= 30) %>%
  arrange(desc(ten_week_hits))

top_19_performers <- top_19_performers %>%
    arrange(desc(ten_week_hits)) %>%
    slice_head(n = 19)

ggplot(top_19_performers, aes(x = fct_reorder(performer, ten_week_hits), y = ten_week_hits, fill = ten_week_hits)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  coord_flip() +  # Flip coordinates for readability
  labs(
    title = "Top 19 Performer with Most Ten-Week Hits (1958-Present)",
    x = "Performer",
    y = "Number of Ten-Week Hits"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    axis.text.y = element_text(size = 12, hjust = 1),  # Increase label size
    axis.text.x = element_text(size = 12),
    axis.title = element_text(size = 14),
    plot.title = element_text(size = 16, face = "bold", hjust = 0.5),
    plot.caption = element_text(size = 10, hjust = 0)
  ) +
  scale_fill_viridis_c() +
  scale_y_continuous(expand = c(0, 0))
```
Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
